---
title: "@TraffLibraries mentions"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: readable
    logo: logo.png
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 4, fig.align = "left")

library(flexdashboard) ; library(knitr) ; library(tidyverse) ; library(rtweet) ; library(tidytext) ; library(reshape2) ; library(igraph) ; library(grid) ; library(ggraph) ; library(lubridate) ; library(DT) ; library(htmltools) ; library(scales)

handle <- "@TraffLibraries"

# ggplot2 theme
theme_x <- function () { 
  theme_minimal(base_size = 14, base_family = "Open Sans") %+replace% 
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(size = 12, face = "bold", hjust = 0),
      plot.subtitle = element_text(size = 10, hjust = 0, margin = margin(9, 0, 9, 0)),
      axis.title = element_text(size = 10, hjust = 1)
    )
}

# get tweets
tweets <- search_tweets(q = paste0(handle, ' OR "Altrincham library" OR "Coppice library" OR "Hale library" OR "Trafford Local Studies" OR "Lostock library" OR "Old Trafford library" OR "Partington library" OR "Sale library" OR "Stretford library" OR "Timperley library" OR "Trafford Town Hall Library" OR "Urmston library" OR "Woodsend library"'),
                          n = 18000,
                          include_rts = FALSE,
                          `-filter` = "replies",
                          retryonratelimit = TRUE,
                          lang = "en") %>%
  mutate(created_at = as.Date(as.POSIXct(created_at)),
         screen_name = factor(screen_name))

# tokenise
tidy_tweets <- tweets %>%
  mutate(text = str_remove_all(text, "&amp;|&lt;|&gt;")) %>% 
  unnest_tokens(word, text, token = "tweets") %>% 
  filter(!word %in% stop_words$word, # remove stop words
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>% # remove numbers
  select(screen_name, word, created_at, is_retweet, hashtags)

# create links
createLink <- function(x) {
  if(is.na(x)){
    return("")
  }else{
    sprintf(paste0('<a href="', URLdecode(x),'" target="_blank">', 
                   substr(x, 1, 25) ,'</a>'))
  }
}

tweets$urls_expanded_url <- lapply(tweets$urls_expanded_url, 
                                     function(x) sapply(x, createLink))
```

Dashboard
=====================================

Row 
-------------------------------------

```{r table}
tweets %>% 
  select(created_at, screen_name, text, retweet_count, favorite_count, urls_expanded_url) %>% 
  datatable(class = 'cell-border stripe',
            colnames = c("Date", "Handle", "Tweet", "Retweets", "Likes", "URL"),
            rownames = FALSE, 
            filter = 'top',
            caption = tags$caption(
            style = 'caption-side: bottom; text-align: center;',
            em(paste0("Tweets mentioning ", handle))),
            width = "100%",
            options = list(
              search = list(regex = TRUE),
              searchHighlight = TRUE,
              dom = 'tip',
              deferRender = TRUE,
              scrollX = TRUE,
              pageLength = 100,
              autoWidth = TRUE,
              columnDefs = list(
                list(className = 'dt-left', targets = "_all"),
                list(width = '70%', targets = 2),
                list(targets = c(0,3,4,5), searchable = FALSE))
              )) %>% 
  formatStyle(columns = 0:4, fontSize = '90%')
```

Row {.tabset .tabset-fade}
-------------------------------------

### Top 10 tweeters
```{r}
tweets %>% 
  count(screen_name, sort = TRUE) %>%
  top_n(10) %>%
  mutate(screen_name = paste0("@", screen_name)) %>% 
  ggplot(aes(x = reorder(screen_name, n), y = n)) +
  geom_col(fill = "#88CCEE") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005),
                     breaks = pretty_breaks()) +
  labs(x = NULL, y = NULL,
       title = "Frequency of tweets by handle",
       subtitle = paste0(handle, " mentions, ", format(min(tweets$created_at), "%d-%m"), " to ", format(max(tweets$created_at),"%d-%m-%Y"))) +
  coord_flip() +
  guides(fill = FALSE) +
  theme_x()
```


### Word frequency
```{r}
tidy_tweets %>% 
  filter(!str_detect(word, "@\\S+")) %>%
  count(word, sort = TRUE) %>% 
  filter(n > 10) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n)) +
  geom_col(fill = "#CC6677") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005),
                     breaks = pretty_breaks()) +
  labs(x = NULL, y = NULL,
       title = "Most common words",
       subtitle = paste0(handle, " mentions, ", format(min(tweets$created_at), "%d-%m"), " to ", format(max(tweets$created_at),"%d-%m-%Y"))) +
  coord_flip() +
  theme_x()
```

### Hashtag frequency
```{r}
tidy_tweets %>% 
  filter(str_detect(word, "^#")) %>% 
  count(word, sort = TRUE) %>% 
  top_n(5) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n)) +
  geom_col(fill = "#DDCC77") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005),
                     breaks = pretty_breaks()) +
  labs(x = NULL, y = NULL,
       title = "Most common hashtags",
       subtitle = paste0(handle, " mentions, ", format(min(tweets$created_at), "%d-%m"), " to ", format(max(tweets$created_at),"%d-%m-%Y"))) +
  coord_flip() +
  theme_x()
```

### Sentiment analysis
```{r}
tidy_tweets %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  top_n(10) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n, fill = sentiment)) +
  geom_col(alpha = 0.7) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_fill_manual(values = c("positive" = "#44AA99", "negative" = "#999999")) + 
  scale_y_continuous(expand = c(0.005, 0.005),
                     breaks = pretty_breaks()) +
  coord_flip() +
  labs(x = NULL, y = NULL, 
       title = "Negative and positive sentiment in top words",
       subtitle = paste0(handle, " mentions, ", format(min(tweets$created_at), "%d-%m"), " to ", format(max(tweets$created_at),"%d-%m-%Y"))) +
  theme_x() +
  theme(legend.position = "none")
```

### Network analysis
```{r}
bigram_graph <- tweets %>%
  mutate(text = str_remove_all(text, "&amp;|&lt;|&gt;"),
         text = str_remove_all(text, "\\s?(f|ht)(tp)(s?)(://)([^\\.]*)[\\.|/](\\S*)")) %>% # remove URLs
  unnest_tokens(word, text, token = "tweets") %>%
  filter(!word %in% stop_words$word, # remove stop words
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]"),
         !str_detect(word, "#\\S+"), # remove hashtags
         !str_detect(word, "@\\S+")) %>% # remove @user
  select(word) %>% 
  unnest_tokens(bigram, word, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  count(word1, word2, sort = TRUE) %>%
  select(word1, word2, n) %>%
  filter(n > 2) %>%
  graph_from_data_frame()

a <- arrow(type = "closed", 
                 length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "#AA4499", size = 5, alpha = 0.7) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  labs(x = NULL, y = NULL, 
       title = "Bigram network graph",
       subtitle = paste0(handle, " mentions, ", format(min(tweets$created_at), "%d-%m"), " to ", format(max(tweets$created_at),"%d-%m-%Y"))) +
  theme_x() +
  theme(panel.grid.major.x = element_blank(),
        axis.text = element_blank())
```

Information {data-icon="fa-info-circle"}
=====================================

This dashboard retrieves tweets over the last 6-9 days that mention Trafford's libraries. Tweets were last retrived at `r paste0(format(Sys.time(), '%H:%M'), ' on ', format(Sys.time(), '%d %B %Y'))` using the following search parameters: '@TraffLibraries  OR "Altrincham library" OR "Coppice library" OR "Hale library" OR "Trafford Local Studies" OR "Lostock library" OR "Old Trafford library" OR "Partington library" OR "Sale library" OR "Stretford library" OR "Timperley library" OR "Trafford Town Hall Library" OR "Urmston library" OR "Woodsend library"'. The [rtweet](https://cran.r-project.org/web/packages/rtweet) [R](https://www.r-project.org/) package was used to retrieve the tweets and [tidytext](https://cran.r-project.org/web/packages/tidytext) to tokenise them for analysis.   

**Please note** that some of the tweets may not be relevant because the search parameters were unable to filter them.